---
title: 'Exploring EV Data'
author: Samuel Adebayo, Jack Kester
output: pdf_document
---

```{r}
#Getting and tidying the electric vehicle data frame
library(tidyverse) 
library(RCurl)
library(emmeans)

electric <- read.csv(text = getURL("https://raw.githubusercontent.com/Stat184-Fall2024/Section002_FP_SamuelAdebayo_JackKester/refs/heads/main/Electric_Vehicle_Population_Data.csv"))

e_clean <- electric %>%
  select(Model.Year:Electric.Range) %>%
  filter(Electric.Vehicle.Type == 
           "Battery Electric Vehicle (BEV)") %>%
  filter(Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility == 
           "Clean Alternative Fuel Vehicle Eligible")
```

```{r}
#Getting the mean for every Make of EV
#Load in Packages
library(ggplot2)
library(esquisse)
library(rvest)
#Getting only the Make and Electric Range
Electric_MakeRange <- e_clean %>%
  select(Make, Electric.Range)

#Make Separate Data Sheets based on Vehicle Make (Does NOT Exist for all Makes)
Tesla_Range <- Electric_MakeRange %>%
  filter(Make == 'TESLA')

Nissan_Range <- Electric_MakeRange %>%
  filter(Make == 'NISSAN')

Chevy_Range <- Electric_MakeRange %>%
  filter(Make == 'CHEVROLET')

Kia_Range <- Electric_MakeRange %>%
  filter(Make == 'KIA')

FIAT_Range <- Electric_MakeRange %>%
  filter(Make == 'FIAT')

Volkswagen_Range <- Electric_MakeRange %>%
  filter(Make == 'VOLKSWAGEN')

BMW_Range <- Electric_MakeRange %>%
  filter(Make == 'BMW')

Hyundai_Range <- Electric_MakeRange %>%
  filter(Make == 'HYUNDAI')

#Create a function that Gives the Average Range with the make table as an input
MakeRange <- function(MakeRange){
  MakeRange %>%
    summarize(
      Make = Make[1],
      AvgRange = mean(Electric.Range, na.rm = TRUE))
  
}

#Create a table that has each Make with its average Range displayed next to it
RangeAvg <- MakeRange(Tesla_Range) %>%
  full_join(MakeRange(Nissan_Range)) %>%
  full_join(MakeRange(Chevy_Range)) %>%
  full_join(MakeRange(Kia_Range)) %>%
  full_join(MakeRange(FIAT_Range)) %>%
  full_join(MakeRange(Volkswagen_Range)) %>%
  full_join(MakeRange(BMW_Range)) %>%
  full_join(MakeRange(Hyundai_Range))

#Now Create a Graph for This data
ggplot(
  data = RangeAvg,
  mapping = aes(
    x = RangeAvg$Make,
    y = RangeAvg$AvgRange,
    fill = RangeAvg$Make
  )
) +
  geom_col() +
  labs(
    x = 'Car Make',
    y = 'Average Electric Range (In Miles)',
    fill = 'Car Make',
    title = 'Average Electric Range by Car Make'
  ) +
  scale_fill_manual(
    values = c('#000000', '#FFF0F0', '#FA0F11', '#0FFC1A', '#A100F1', '#00FAFA', '#000AFA', '#FA11FC')
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```
Research Questions:
Q1: Is there a difference in average electric range between the Tesla models? 



```{r}

tesla <- e_clean[e_clean$Make == "TESLA",]

boxplot(Electric.Range~Model, data = tesla)

tesla_fit <- aov(Electric.Range~Model, data = tesla)

summary(tesla_fit)

plot(fitted(tesla_fit), rstandard(tesla_fit), 
     xlab = "Fitted Values", ylab = "Standardized Residuals")

qqnorm(rstandard(tesla_fit))
qqline(rstandard(tesla_fit))



```
Research Questions:
Is there a difference in average electric range between the Tesla models? 

Response:

Q1: Anova Hypothesis Test (At 5% sig level)
(1) $H_0: \mu_1=\mu_2=\mu_3$
$H_1: Some\ \mu_i \neq \mu_k$

(2)
```
               Df   Sum Sq Mean Sq F value Pr(>F)    
Model           4  6903021 1725755    1347 <2e-16 ***
Residuals   25299 32408761    1281 
```

(3)
Validate Assumptions
Independence: Cars aren't randomly assigned Models because they are manufactured specifically as that model, however we can assume cars were randomly chosen to be tested so we can validate independence

Normality: Normality tails off on the sides, however slightly follows a normal distribution

Constant Variance: Variance does not randomly disperse around 0 so cannot really validate variance.

Assumptions cannot be fully validated, however, to respond properly to the question we will continue

(4)
P-Value = P(F >= 1347) = [(<2e-16) < 0.05 (significance level)]

Because our p-value is small (<5%), we can conclude that at least one mean electric range is different than the others.
